Notes on generating initial models for mapping into {\tt MAESTRO}

### 2008-10-24 docs/initial_models/initial_models.tex
remove extraneous \end{document}
define a macro for the figure path to allow this to be imbedded in other documents
switch over to book format for the master document -- this makes the sectioning more natural

\documentclass[11pt]{article}

\tolerance=600

\usepackage{amsmath,color,epsfig,epsf}

% Margins
\usepackage[lmargin=0.5in,rmargin=1.5in,tmargin=0.5in,bmargin=0.5in]{geometry}

\newcommand{\sfrac}[2]{\mathchoice
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
  {#1\!/#2}}

\def\half  {\frac{1}{2}}
\newcommand{\myhalf}{\sfrac{1}{2}}

\def\dt    {\Delta t}
\def\kth   {k_{\rm th}}
\def\init  {\rm init}
\def\model {\rm model}
\def\raw   {\rm raw}

\def\edge  {\rm EDGE}
\def\mac   {\rm MAC}
\def\trans {\rm TRANS}

\def\eb    {{\bf e}}
\def\fb    {{\bf f}}
\def\ib    {{\bf i}}
\def\Ub    {{\bf U}}

\def\Ubt   {\widetilde{\bf U}}
\def\ut    {\tilde{u}}
\def\vt    {\tilde{v}}
\def\wt    {\tilde{w}}

\title{Notes on generating initial models for mapping into {\tt MAESTRO}}

\setlength{\marginparwidth}{1.0in}
\newcommand{\MarginPar}[1]{\marginpar{%
\vskip-\baselineskip %raise the marginpar a bit
\raggedright\tiny\sffamily
\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}

\begin{document}

\maketitle
\tableofcontents

\clearpage

@@

\subsection{Multilevel Plane-Parallel}
We do the same thing as the single-level case except that we generate
initial data at the non-finest levels by using linear interpolation
from the two nearest model points (see Figure \ref{Fig:Multilevel
  Initialization}).  Note that the initial data is not in HSE nor is
it thermodynamically consistent on non-finest cells.  We will enforce
these on the base state and full state later.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tpb]
\begin{center}
\epsffile{./multilevel_init.eps}
\caption{Multilevel Initialization.  The data from the initial model
  is represented by the dots on the right.  The initial data at the
  finest level is represented by the circles.  The initial data at
  non-finest levels is represented by the squares.  We copy data from
  the dots directly into the circles.  We use linear interpolation
  using the two nearest points to copy data from the dots to the
  squares.}
\label{Fig:Multilevel Initialization}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsubsection{Coarse-Fine HSE Discretization}\label{Sec:Coarse-Fine HSE Discretization}
Integrating the HSE discretization upward, when we encounter a change
from coarse to fine, here is an equation to integrate up to the
coarse-fine interface ($l$ is the ``finer'' level, $l-1$ is the
``coarser'' level)
\begin{equation}
\frac{p_{r-\myhalf}^l - p_{\sfrac{r}{2}-1}^{l-1}}{\Delta r^{l-1}/2} = \frac{\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}}{2}g ~~~ \rightarrow ~~~ p_{r-\myhalf}^l = p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^{l-1} g}{4}\left(\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}\right).
\end{equation}
Here is an equation to integrate from the coarse-fine interface up to
the cell center
\begin{equation}
\frac{p_r^l - p_{r-\myhalf}^l}{\Delta r^l/2} = \frac{\rho_r^l + \rho_{r-\myhalf}^l}{2}g ~~~ \rightarrow ~~~ p_r^l = p_{r-\myhalf}^l + \frac{\Delta r^l g}{4}\left(\rho_r^l + \rho_{r-\myhalf}^l\right).
\end{equation}
Combining equations gives
\begin{equation}
p_r^l = p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^{l-1} g}{4}\left(\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}\right) + \frac{\Delta r^l g}{4}\left(\rho_r^l + \rho_{r-\myhalf}^l\right).
\end{equation}
We can simplify using
\begin{equation}
\Delta r^{l-1} = 2\Delta r^l,
\end{equation}
\begin{equation}
\rho_{r-\myhalf}^l = \frac{2}{3}\rho_r^l + \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1}.
\end{equation}
Simplifying
\begin{eqnarray}
p_r^l &=& p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^l g}{2}\left(\frac{2}{3}\rho_r^l + \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1} + \rho_{\sfrac{r}{2}-1}^{l-1}\right) + \frac{\Delta r^l g}{4}\left(\rho_r^l + \frac{2}{3}\rho_r^l + \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1}\right) \nonumber \\
&=& p_{\sfrac{r}{2}-1}^{l-1} + \frac{3\Delta r^l g}{4}\left(\rho_{\sfrac{r}{2}-1}^{l-1} + \rho_r^l\right).\label{Coarse-Fine Stencil}
\end{eqnarray}
When we encounter a change from fine to coarse, analogously
\begin{equation}
p_{(r+1)/2}^{l-1} = p_{r}^l + \frac{3\Delta r^l g}{4}\left(\rho_{r}^l+\rho_{(r+1)/2}^{l-1}\right).\label{Fine-Coarse Stencil}
\end{equation}

\end{document}

>>>>

@@

\subsection{Multilevel Plane-Parallel}
We do the same thing as the single-level case except that we generate
initial data at the non-finest levels by using linear interpolation
from the two nearest model points (see Figure \ref{Fig:Multilevel
  Initialization}).  Note that the initial data is not in HSE nor is
it thermodynamically consistent on non-finest cells.  We will enforce
these on the base state and full state later.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tpb]
\begin{center}
\epsffile{\figpath/multilevel_init.eps}
\caption{Multilevel Initialization.  The data from the initial model
  is represented by the dots on the right.  The initial data at the
  finest level is represented by the circles.  The initial data at
  non-finest levels is represented by the squares.  We copy data from
  the dots directly into the circles.  We use linear interpolation
  using the two nearest points to copy data from the dots to the
  squares.}
\label{Fig:Multilevel Initialization}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Coarse-Fine HSE Discretization}\label{Sec:Coarse-Fine HSE Discretization}
Integrating the HSE discretization upward, when we encounter a change
from coarse to fine, here is an equation to integrate up to the
coarse-fine interface ($l$ is the ``finer'' level, $l-1$ is the
``coarser'' level)
\begin{equation}
\frac{p_{r-\myhalf}^l - p_{\sfrac{r}{2}-1}^{l-1}}{\Delta r^{l-1}/2} = \frac{\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}}{2}g ~~~ \rightarrow ~~~ p_{r-\myhalf}^l = p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^{l-1} g}{4}\left(\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}\right).
\end{equation}
Here is an equation to integrate from the coarse-fine interface up to
the cell center
\begin{equation}
\frac{p_r^l - p_{r-\myhalf}^l}{\Delta r^l/2} = \frac{\rho_r^l + \rho_{r-\myhalf}^l}{2}g ~~~ \rightarrow ~~~ p_r^l = p_{r-\myhalf}^l + \frac{\Delta r^l g}{4}\left(\rho_r^l + \rho_{r-\myhalf}^l\right).
\end{equation}
Combining equations gives
\begin{equation}
p_r^l = p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^{l-1} g}{4}\left(\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}\right) + \frac{\Delta r^l g}{4}\left(\rho_r^l + \rho_{r-\myhalf}^l\right).
\end{equation}
We can simplify using
\begin{equation}
\Delta r^{l-1} = 2\Delta r^l,
\end{equation}
\begin{equation}
\rho_{r-\myhalf}^l = \frac{2}{3}\rho_r^l + \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1}.
\end{equation}
Simplifying
\begin{eqnarray}
p_r^l &=& p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^l g}{2}\left(\frac{2}{3}\rho_r^l + \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1} + \rho_{\sfrac{r}{2}-1}^{l-1}\right) + \frac{\Delta r^l g}{4}\left(\rho_r^l + \frac{2}{3}\rho_r^l + \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1}\right) \nonumber \\
&=& p_{\sfrac{r}{2}-1}^{l-1} + \frac{3\Delta r^l g}{4}\left(\rho_{\sfrac{r}{2}-1}^{l-1} + \rho_r^l\right).\label{Coarse-Fine Stencil}
\end{eqnarray}
When we encounter a change from fine to coarse, analogously
\begin{equation}
p_{(r+1)/2}^{l-1} = p_{r}^l + \frac{3\Delta r^l g}{4}\left(\rho_{r}^l+\rho_{(r+1)/2}^{l-1}\right).\label{Fine-Coarse Stencil}
\end{equation}
